Spin canonical divisors of log stacky curves

### 2015-06-21 stacky_write_up/spin_cring.tex
added inductive lemma for genus > 0

\ssec{Inductive Theorem: Increasing the Number of Ramified Points}
\label{ssec:g_0_ram_pts}

@@

\begin{prop}
\label{prop:sat_three_induction}
Let $\sx \rightarrow \sx'$ be a birational map of tame, separably
rooted log stacky curves $R \subset R'$
\todo{Should switch either $R$ and $R'$ or $\sx$ and $\sx'$ to make prime notation consistent (note that inclusion is reversed since this map adds a stacky point to $\sx$; note that we should also say that this map is only ramified at $P$)} 
be the corresponding
inclusion of canonical rings. Let $R_{D'} = R' = k[x]/I'= k[x_3, x_4
, x_5,\ldots, x_m]/I'$ satisfy $sat(Eff(D')) = 3.$ Let $D = \frac{1}{
3}P$\todo{Peter: Do you mean to have $D=\frac{1}{3}P+D'$ or is it correct as written?}.
\todo{Peter: We can generalize to arbitrary genus with the condition $\deg(D)>2g-2$ meaning $h^0(X,K-D)=0$ so Riemann-Roch gives us everything we wanted from $\mathbb{P}^1$.}

\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3,
		4,5\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ and$ y_5$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3,y_4,y_5]$ with $grlex$ so that $y_3 \prec 
		y_4 \prec y_5$ 
		\todo{This order isn't quite right, we need $y_4^2 < y_5y_3$.}
		and equip the ring $k[y_3,y_4,y_5,x]$ with the block 
		order so that $R = k[y_3,y_4,y_5,x]/I$. Let,
		\begin{align*}
			J &= gin_\prec(I')k[x,y_3,y_4,y_5] \\
			&+\langle y_i x_j \mid 4 \leq i \leq 5, 3 \leq j \leq m\rangle \\
			&+\langle y_iy_k \mid 4 \leq i \leq j \leq 5.\rangle 
		\end{align*}
			Then, $gin_\prec(I) = J$.
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{prop}

@@

\begin{align}
\label{eqn:add_one_generator}
	\begin{split}
		&\langle y_3^ax_3^b x_4^\epsilon x_5^{\epsilon'} \mid a \geq 0, b 
		\geq 0,(\epsilon,\epsilon') \in \{(0,0),(0,1),(1,0)\} \rangle \\
		&+\langle y_3^ay_4,y_3^by_5 \mid a \geq 0, b \geq 0 \rangle
	\end{split}
\end{align}
\todo{You need to say explicitly somewhere that $x_i$ lies in degree $i$; also for the higher genus case, we should replace $x_3, x_4, x_5$ with $x_2, x_3$ and everything should work the same.}

@@

\section{Genus positive}
We haven't done this.

>>>>

\ssec{Inductive Theorem: Increasing the Number of Ramified Points for a 3-saturated divisor}
\label{ssec:g_0_ram_pts}

@@

\begin{prop}
\label{prop:sat_three_induction}
\todo{Make explicit assumptions on genus.}
\todo{Should switch either $R$ and $R'$ or $\sx$ and $\sx'$ to make prime notation consistent (note that inclusion is reversed since this map adds a stacky point to $\sx$; note that we should also say that this map is only ramified at $P$), Aaron: you're right, I wrote the inclusion on rings in the wrong direction. Actually, the map isn't ramified at $P$, it's just defined away from $P$. I decided to remove this from the proposition though.} Let $R_{L'} = R' = k[x]/I'= k[x_3, x_4
, x_5,\ldots, x_m]/I'$ satisfy $sat(Eff(D')) = 3.$ Let $L = L' + \frac{1}{
3}P$, and let $R = R_L$. Suppose $\deg x_i = i$ for $i \in \{3, 4, 5\}$ and that the ordering on $k[x]$ satisfies 
\begin{align*}
	\ord_{x_3}(f) < \ord_{x_3}(g) \implies f \prec g.
\end{align*}
Then, the following statements hold.

\todo{Peter: Do you mean to have $D=\frac{1}{3}P+D'$ or is it correct as written?}.
\todo{Peter: We can generalize to arbitrary genus with the condition $\deg(D)>2g-2$ meaning $h^0(X,K-D)=0$ so Riemann-Roch gives us everything we wanted from $\mathbb{P}^1$.}

\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3,
		4,5\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ and$ y_5$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4, y_5]$ with $grlex$ so that $y_3 \prec 
		y_4 \prec y_5$
		and equip the ring $k[y_3, y_4, y_5, x]$ with the block 
		order so that $R = k[y_3, y_4, y_5, x]/I$. Let,
		\begin{align*}
			J &= gin_\prec(I')k[x,y_3,y_4,y_5] \\
			&+\langle y_i x_j \mid 4 \leq i \leq 5, 3 \leq j \leq m\rangle \\
			&+\langle y_iy_k \mid 4 \leq i \leq j \leq 5.\rangle 
		\end{align*}
			Then, $gin_\prec(I) = J$.
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{prop}

@@

\begin{align}
\label{eqn:add_one_generator}
	\begin{split}
		&\langle y_3^ax_3^b x_4^\epsilon x_5^{\epsilon'} \mid a \geq 0, b 
		\geq 0,(\epsilon,\epsilon') \in \{(0,0),(0,1),(1,0)\} \rangle \\
		&+\langle y_3^ay_4,y_3^by_5 \mid a \geq 0, b \geq 0 \rangle
	\end{split}
\end{align}

@@

\section{Genus positive}

\ssec{Inductive Theorem: Increasing the Number of Ramified Points for a 2-Saturated Divisor}

\begin{prop}
\label{prop:sat_two_induction}
Let $(\sx, \Delta, L)$ and $(\sx', \Delta, L')$ be log stacky spin curves. Let $\sx \rightarrow \sx'$ be a birational map of tame, separably rooted log stacky curves so that $\sx$ has signature $\sigma = (g;e_1,\ldots, e_r;\delta)$ with $g > 0,$ and, if $g = 1$ then $\deg 3L \geq 2$.
Let $R_{L'} = R' = k[x]/I'= k[x_2, x_3
, x_5,\ldots, x_m]/I'$ and let $L = L' + \frac{1}{
3}P.$ Suppose $L$ and $L'$ satisfy $sat(Eff(L')) = sat(Eff(L)) = 2.$ Suppose $\deg x_i = i$ for $i \in \{2, 3\}$ and that the ordering on $k[x]$ satisfies
\begin{align*}
	\ord_{x_2}(f) < \ord_{x_2}(g) \implies f \prec g.
\end{align*}
Then, the following statements hold.
\begin{enumerate}
	\item[(a)] General elements  $y_i \in H^0(\sx,iL)$ for $i \in \{3, 4\}$ satisfy $-\ord_P(y_i) = 1$ and any such choice of elements $y
		_3,y_4,$ minimally generate $R$ over $R'$.
	\item[(b)] Equip $k[y_3, y_4]$ with $grlex$ so that $y_3 \prec 
		y_4$
		and equip the ring $k[y_3, y_4, x]$ with the block 
		order so that $R = k[y_3, y_4, x]/I$. Let,
		\begin{align*}
			J &= gin_\prec(I')k[x, y_3, y_4] \\
			&+\langle y_4 x_j \mid 3 \leq j \leq m \rangle \\
			&+\langle y_4^2.\rangle
		\end{align*}
			Then, $gin_\prec(I) = J$.
	\item[(c)] Any set of minimal generators for $I'$ together with 
		any set of relations with leading terms as in (b) minimally 
		generate $I$.
\end{enumerate}
\end{prop}

\begin{proof}
The assumptions on $g,\sigma$ imply $H^0(\sx, 3L), H^0(\sx, 4L)$ are both base point free: If $g \geq 2$ then $\deg 3L > 2g-1$ and $\deg 4L > 2g-1$, so $H^0(\sx, 3L)$ and $H^0(\sx, 4L)$ are base point free. If $g = 1$, we assume $\deg 3L \geq 2 > 2g-1$, so we also have $\deg 4L \geq 2 > 2g -1,$ so again $H^0(\sx, 3L)$ and $H^0(\sx, 4L)$ are base point free. Therefore, general elements $y_3$ and $y_4$ satisfy $-\ord
_P(y_i) = 1$ by Riemann Roch, and the assumptions on $\sigma$. \todo{Understand why/whether we need 
to cite lemma 5.4.7.}

In a fashion analogous to the proof of Proposition 
~\ref{prop:sat_three_induction}, we see 
the following is a $k$ basis for $R$ over $R'$:

\begin{align}
\label{eqn:sat_two_add_generator}
	\begin{split}
		&\langle y_3^ax_3^b x_4^\epsilon \mid a \geq 0, b 
		\geq 0, \epsilon \in \{0, 1\}\rangle \\
		&+\langle y_3^ay_4 \mid a \geq 0 \rangle,
	\end{split}
\end{align}

\noindent completing part (a).

To show part (b), note that the generators in
~\ref{eqn:sat_two_add_generator} are precisely a set of monomials which
generate $k[x,y]/J$ over $k$. So, to show $J = in_\prec(I)$, it
suffices to show that all generators of $J$ lie in $in_\prec(I)$.

This follows, since there exist constants $A_j \in k$ for $3 \leq j \leq m,$ and a constnant $B \in k$ and
elements $w_{i,j} \in R'$ so that the following linear combinations
of elements lie in $R'$.

\begin{align*}
	y_4x_j &- A_j y_3 w_j & \text{ so that } 3 \leq j \leq m \text{ and } \deg w_j = \deg x_j + 1\\
	y_4^2 &+ B y_3^2 x_2 \\
\end{align*}
